\documentstyle{article}
\nofiles
\pagestyle{empty}
\begin{document}
\newfont{\iching}{iching60}

\newcount\hexagram
\newcount\hexachar
\newcount\changeline
\newcount\random

\random=1234
\newcount\clchar
\hexagram=\random
\divide\hexagram by 64 
\advance\hexagram by 1
\multiply\hexagram by 64
\advance\hexagram by -\random

\changeline=\random
\divide\changeline by 6
\advance\changeline by 1
\multiply\changeline by 6
\advance\changeline by -\random



\newcommand{\interpret}[9]{
{\bf #1}\hfill{\iching\symbol\hexachar}\hfill{\bf #2}

\vfill
\begin{center}{\iching\symbol\clchar\symbol\hexagram}\end{center}

\vfill

{\sf Interpretation of hexagram \number\hexagram}

{\it #3}

\vfill
\begin{center}{\iching\symbol{0}}\end{center}

\vfill
{\sf Interpretation of change line \number\changeline}

\ifcase \changeline 
\or {\it #4}
\or {\it #5}
\or {\it #6}
\or {\it #7}
\or {\it #8}
\or {\it #9}
\fi		}

\clchar=\changeline
\advance\clchar by 200
\hexachar=\hexagram
\advance\hexachar by 64
\ifcase \hexagram
\or \interpret{CH'IEN}{The Origin}
	{The origin. Continuing through. Harmonious. Correct.}
	{Do not use the diving dragon.}
	{The dragon is seen in the field. It is helpful to see the great man.}
	{This wise man is active all day long, but in the evening he is watchful.
		There will be great difficulty.}
	{He is bouncing up and down in stormy waters. The sage says that he sometimes leaps up and sometimes
   		down. There is nothing wrong in this as long as evil is avoided.}
	{The dragon is flying in the heavens. It is good to see the great man.}
	{A group of headless dragons appears. There will be good fortune.}
\or \interpret{K'UN}{Success}
	{Great success. A wise man is virtuous and proper, like a strong but gentle mare. When he is faced with a
		task he can only succeed if he takes things slowly and carefully, not trying to be first. If he is
		happy to follow others, all will be well. It is important to have friends in the south and west,
		but they can be lost or forgotten if they are in the north and east. A wise man must remember to
		draw strength from proper conduct, and by doing this he will be fortunate.}
	{He is walking on newly formed frost. Freezing conditions are on the way.}
	{Square, straight and strong. Everthing will work out well without any effort.}
	{He does not reveal his virtue for all to see, but nurtures it within himself. He has the opportunity
   		of serving the Emperor and will perform his tasks with great success.}
	{A sack has been sealed up so that nothing can flow in or
	out. There will be no blame and no praise. What is done is done.}
	{Imperial yellow undergarments. There will be excellent fortune.}
	{There are dragons involved in a ferocious fight in the desert. The blood that flows from their wounds
   		is Imperial yellow.}
\or \interpret{CHUN}{Birth Pangs}
	{Birth pangs. He will succeed if he behaves properly. Beware of starting a new venture without giving
	it sufficient thought. It will be useful to enlist the help of an experienced person.}
	{The way ahead is blocked. Remain firm. Act sensibly. Listen to the advice of someone more experienced.}
	{The horses are outside the village but are not allowed inside. The people think they are robbers.
   		They have come to collect the bride. The bride remains firm and rejects the union. Ten years later
		she marries and has children.}
	{The hunter, stalking a deer, is drawn into the forest. He failed to ask the forester for help and is
   		now lost. The wise man can see the dangers which lie ahead. He know it would be foolish to
		continue, so decides to give up the chase.}
	{The woman has lost control of her carriage and the horses are running away with her. The man who
   		wants to marry her is nearby and she calls for his help. To advance will bring good fortune.
		All will turn out well.}
	{He is finding it difficult to grant the favours people expect of him. Wise behaviour brings success
   		when dealing with minor affairs. Any attempts to deal with major issues will bring misfortune.}
	{The man is in complete disarray since his chariot horses have gone into full retreat. He cannot hold
   		back his tears and begins to weep.}
\or \interpret{MENG}{Rebellious Youth}
	{Rebellious youth. I do not look for the rebellious and ignorant child; it is he who looks for me. When
   		he asks me for an oracle I will answer him. But when he asks me a second and third time he becomes
		a nuisance. I refuse to answer his requests. It is helpful to be consistent and well behaved.}
	{It is good to use punishment in order to banish ignorance. This will free the mind from its bondage.
		Too much punishment brings about humiliation.}
	{Be patient with the ignorant, since this will bring good fortune. It is good to listen to the
		requests of women.  A son is able to run his family.}
	{Do not become involved with a woman who will easily give herself to a man with money. Such
		behaviour will bring misfortune.}
	{It is sordid to be constrained by ignorance. It will bring regret.}
	{The innocent child. There will be good fortune.}
	{Punishing the ignorant child will do no good. It is better to protect him.}
\or \interpret{HSU}{Patience}
	{Patience. Confidence and sicerity will bring prosperity. Virtue will encourage good fortune. There is
		much to be gained by crossing the great river.}
	{He is waiting in open country. A lot can be gained by using that which endures.}
	{He is waiting on the sand. He will have to tolerate gossip, but in the end he will be very fortunate.}
	{He is sitting out in the mud. This allows the enemy to arrive.}
	{He is waiting in a place of blood. He must escape from the pit.}
	{He is waiting amongst the food and drink. Good fortune will come through virtue.}
	{He goes into the pit. Three guests arrive who have not been invited. Treat them well and in the
   		end there will be good fortune.}
\or \interpret{SUNG}{Contention}
	{Contention. He is confident in the face of opposition. Respectable behaviour will bring good fortune.
		If he persists to the bitter end there will be misfortune. It is very helpful to see the great man.
		It is not a good idea to cross the great river.}
	{He never pursues the issue. He has to tolerate gossip. Eventually he will be lucky.}
	{He does not have the strength to face the challenge. If he decides to run back home to the three
   		hundred families of his city he will bring no harm upon himself.}
	{He stays in the correct place. There are great difficulties at the moment but everything will work
		out well in the end. If he should happen to be in the king's service he will not achieve all that
		he hoped for.}
	{He does not have the strength to face up to the challenge. He retreats and studies the way fate
		deals with people. He changes his attitude and finds peace in virtuous behaviour. There will
		be good fortune.}
	{By attending in the law court extremely good fortune will follow.}
	{The leather belt of status may be given to him but it will be taken away from him three times in
		one morning.}
\or \interpret{SHIH}{The Army}
	{The army. Everything is correct. Nothing will go wrong is the leader is wise and experienced.}
	{The army marches off after receiving its orders. It will be disasterous if these orders are wrong
		or not well thought out.}
	{There is good fortune since the commander is at the centre of the army. He did not make any mistakes.
		The king honours him three times.}
	{The army has incompetent leaders. There will be misfortune.}
	{The army retreats. There are no mistakes.}
	{There are many different species of birds in the field. Now is the time to sieze them to avoid
		mistakes. The eldest son is the leader of the army. Although a younger son acts wisely he will
		certainly fail in the role of commander because evil will befall him.}
	{The great prince in is complete charge. He chooses men to be rulers of countries, and others he
		appoints as the heads of powerful families. He does not give any of these appointments to weaker men.}
\or \interpret{PI}{Unity}
	{Unity. There will be good fortune. When he asks the oracle about his fortune he is
		reassured. Those who are concerned will gather together at the right time. Those
		who arrive late will be unlucky.}
	{Confident actions result in unity. Confidence is like an overflowing cup which will
		eventually bring a variety of good fortune.}
	{Unity develops from the inner self. Good fortune will result from righteous behaviour.}
	{He looks to unscrupulous people for unity.}
	{Unity is sought with those outside his close circle. Wise behaviour will bring
		good fortune.}
	{The need for unity is becoming obvious. The king hunts the wildfowl on three sides
		only; he leaves the fourth side open for their escape. The local people were not
		instructed properly. There will be good fortune.}
	{There is nobody to lead the united group. It is an evil time.}
\or \interpret{HSIAO CH'U}{Holding Back the Less Able}
	{Holding back the less able. Progress and prosperity will develop. Clouds approach from
		our lands to the west but they do not bring rain.}
	{He returns to the true path. Why should he be blamed for this? There will be good fortune.}
	{He is lured back. There will be good fortune.}
	{The spokes in the carriage wheel are broken. The husband and wife turn away from each other.}
	{Have confidence. The risk of bloodshed passes and fears disappear. There will be no mistakes.}
	{The feeling of confidence is like the security of being bound together. He has ample
		resources close at hand and can draw his neighbours to him.}
	{When the rain falls everything stops, but virtue continues to flourish. The woman who
		behaves wisely is still in danger. It is nearly full moon. The man who ventures
		forth at this time will cause or create evil.}
\or \interpret{LI}{Walking Carefully}
	{Walking carefully. Although the man treads on the tiger's tail he is not bitten. There is success.}
	{He is walking along his normal path. As long as he continues to do so there will be no problems.}
	{He walks on the path which is smooth and easy. The peaceful hermet will be lucky,
		if he acts honourably.}
	{The one-eyed man can see; the lame mane can walk; but the man who treads on the tail of
		the tiger is bitten and will have bad luck. The army officer behaves in this way for the
		benefit of his great ruler.}
	{He walks on the tail of the tiger. He will eventually be lucky if he is careful
		and cautious.}
	{He walks resolutely. Although his behaviour is correct he is surrounded by danger.}
	{Look at the path which is taken and examine the resulting good luck. When everything
		is fulfilled there will be great fortune.}
\or \interpret{T'AI}{Benevolence}
	{Benevolence. The lesser has disappeared and the greater is yet to come. There will be
		success and good fortune.}
	{When grass is pulled up it brings other stalks with it. Persevere. There will be good fortune.}
	{He bears with that which is uncontrolled. He is confident and crosses the river without
		using a boat. He does not forget the things which are far away from him and does not
		worry about friends. In this way he will keep to the middle path.}
	{Peace cannot exist without disruption. He cannot go without coming back. The man who is
		steadfast will not be blamed. He is not sorry but confident. He enjoys his good fortune.}
	{There is disruption and bustle. He does not rely on his own finances but turns to his
		neighbours. He puts his confidence in them. He does not withdraw from them.}
	{The Emperor Yi gives his younger sister in marriage. This blessing brings great fortune.}
	{The city wall collapses into the moat. Do not summon the army. He tells the people
		of his own country what is happening. Nevertheless, proper behaviour still causes distress.}
\or \interpret{P'I}{Obstruction}
	{Obstruction. Evil men block the path of progress. Events turn out badly for the wise man even
		when he acts correctly. The great are leaving and the less important ones are arriving.}
	{Newly pulled-up grass will bring up other stalks which are attached to it. Firm and correct
		actions will result in progress and good fortune.}
	{He is guided by patience and endurance The lesser man will be lucky and the obstacles
		which lie in the great man's path will lead to success.}
	{He bears with his shame.}
	{He is wise to follow the demands which are made by destiny. When his companions come
		they rejoice in his blessings.}
	{The obstruction is removed. The great man is fortunate. He should take care and be on
		his guard against destruction. He should act as though he is bound to a mulberry bush.}
	{At first there are obstacles. They have now been cast aside and there is rejoicing.}
\or \interpret{T'UNG JEN}{Companions}
	{Companions. Friends will be found in the remote countryside. He will be successful. It is helpful to
		travel across the great river. The wise man will reap rewards if he behaves correctly.}
	{Friendships should be developed with people at the door. There will be no mistakes.}
	{Friendships formed within the family bring humiliation.}
	{He hides the weapons in the grass on top of a hill. He does not attack for three years.}
	{He scales the defences of the city but is unable to attack. There will be good fortune.}
	{At first his companions cry out, then they weep and finally they laugh. The great army
		conquers all in its path. They meet together.}
	{Good friends are to be found in the countryside. There is nothing to regret.}
\or \interpret{TA YU}{Many Possessions}
	{Many possessions. It bodes well for him to have numerous possessions and great success.}
	{He is wise not to approach dangerous areas or people. If he understands the way things
		work there will be no mistakes.}
	{The heavily loaded wagon will be safe whichever way it goes.}
	{The duke offers his labours to the son of Heaven. A lesser man could not do this.}
	{He does not use up his strength. There will be no mistakes.}
	{His self-confidence sets an example for others. Such dignity will bring good fortune.}
	{He is fortunate since Heaven is his personal protector. There is nothing wrong in this.}
\or \interpret{CH'IEN}{Modesty}
	{Modesty. The wise man will eventually succeed in everything.}
	{The wise man carries out his duties with true modesty. Because of his unassuming nature he is
		able to cross the great river. There will be good fortune.}
	{Modesty speaks for itself. Good luck will come from proper conduct.}
	{The superior man succeeds because he is a person of merit and honesty. Thus he has good fortune.}
	{Modesty is an admirable quality.}
	{He is not considered rich by his neighbour's standards. The time is right for invasion.
		Everything he does will work out to his advantage.}
	{Humility has made itself known. It is the right time to attack both city and the country.}
\or \interpret{YU}{Enthusiasm}
	{Enthusiasm. The time is right to appoint princes and give orders for the army to approach.}
	{He shows enthusiasm. There will be misfortune.}
	{He is as solid as a rock. He does not wait until the end of the day. His wise
		behaviour brings good luck.}
	{He raises his head with enthusiasm. This only brings regret. Delays will result in sorrow.}
	{The source of enthusiasm is also the bringer of success. It is better for him to gather
		friends around him than to spend his time worrying.}
	{Although he is very ill he perseveres and does not die.}
	{His enthusiasm is misguided. As long as he has his strength to change direction
		there will be no problems.}
\or \interpret{SUI}{According or Agreeing With}
	{Agreeing with. Great success. It is helpful to behave with propriety. There will be no mistakes.}
	{The official changes. He is fortunate because of his wise behaviour. He is credited for
		going out through the gate to join his friends.}
	{He becomes attached to the young boy and rejects the older and wiser man.}
	{He becomes attached to the older and wiser man and rejects the small boy. By following such a course
		he will find what he is looking for. He will benefit by following the righteous path.}
	{Dogged perseverance is bad but agreeing with others brings fortune. His confidence
		in the path ahead is founded on true understanding. Can you find anything to blame
		him about is that?}
	{He is confident in that which is good. There will be good fortune.}
	{He is bound down by strong links. The king brings fortune to the Western Mountain.}
\or \interpret{KU}{Decay}
	{Decay. Crossing the great river will bring good fortune. Check everything carefully three
		days beforehand and three days afterwards.}
	{The son sorts out the chaos which is caused by his father. As long as he is competent
		he can solve these problems without difficulties. Although there are immense problems
		everything will work out well in the end.}
	{The son has to sort out the chaos caused by his mother. He should not be overzealous in this.}
	{The son sorts out the chaos which his father caused. There is very little to repent
		and he will make no mistakes.}
	{He tolerates the chaos caused by his father. If he continues in this way he will
		be humiliated.}
	{The son is praised for sorting out the chaos which his father caused.}
	{He does not work for the king or his lord since he wants to serve a higher ideal.}
\or \interpret{LIN}{To Draw Near}
	{To draw near. This is a time of great success which is helped by good behaviour. The eighth
		month is an unlucky time.}
	{He draws close to his friends. He will have good luck if he behaves wisely.}
	{It is to his advantage to draw close to his friends. Good luck will come.}
	{He will not suffer setbacks if he does draw close. He will not make any mistakes if
		he thinks his activities through properly.}
	{Nothing will go wrong when he arrives in great style.}
	{Like an eminent leader, he approaches wisely. He will have good fortune.}
	{No mistakes will be made if he approaches with honesty and generosity. He will have
		good fortune.}
\or \interpret{KUAN}{Examine}
	{Examine. He washes his hands in preparation but has not yet made the sacrifice. He inspires
		confidence in others because of his dignified appearance.}
	{He looks at things like a boy. This behaviour is acceptable in lesser men but not in a
		superior man.}
	{He looks out of the door. This kind of behaviour is acceptable in a woman.}
	{He examines his life so that he can decide whether to go forward or retreat.}
	{It is useful to be the king's guest since he can consider the achievements of the kingdom.}
	{He examines his life. The wise man does not make mistakes.}
	{He examines his character to see if he is wise. He will not lapse into error.}
\or \interpret{SHIH HO}{Biting Through}
	{Biting through. Success. It is helpful to use the law.}
	{His feet are trapped and his toes are cut off. There will be no mistakes.}
	{He sinks his teeth into tender flesh and bites off the nose. There will be no mistakes.}
	{He bites on tough meat and it is poisoned. This is not a mistake but it is a cause
		for minor regrets.}
	{He bites on hard, bony meat. He receives the golden arrow. He will bennefit from being
		firm and will recognize the dangers. There will be good fortune.}
	{He discovers a small piece of gold when he bites on dried meat. He should be firm and
		recognize the dangers in order to avoid mistakes.}
	{He is trapped in a wooden collar and has no ears. There will be bad fortune.}
\or \interpret{PI}{To Adorn}
	{To adorn. Success. There is little to be gained in giving him permission to undertake anything.}
	{He adorns his toes with finery. He steps out of the carriage and continues on foot.}
	{He adorns his beard.}
	{His body is bejewelled as though sprinkled with dew. Constancy will bring good luck.}
	{His body is adorned so that he looks as though he is shining with light. He travels on a white, winged
		horse. He is not a thief but is looking for someone to marry.}
	{His clothes and body are decorated with finery and he travels through hills and gardens. His roll
		of silk is poor and small. This may cause distree but in the end there will be good fortune.}
	{He is adorned only in white. There will be no mistakes.}
\or \interpret{PO}{Peeling or Splitting}
	{Peeling or splitting. There is nothing to be gained by moving anywhere.}
	{He breaks the leg of the bed. It is bad to destroy anything which is good.}
	{He breaks the frame of the bed. It is bad to destroy things which are good and useful.}
	{He is the one who breaks things. This is not a mistake.}
	{He breaks both the bed and the skin. This is bad.}
	{He is kind to the people of the palace, treating them like a shoal of fish. There are many advantages
		in this.}
	{He is like a ripe fruit ready to be eaten. The wise man travels in a special carriage. A lesser man's
		home breaks up.}
\or \interpret{FU}{Return}
	{Return. Success. The man travelling to and fro will not be disrupted. His friends come to him and there is
		nothing wrong with this. It is in his way to come and go. He returns on the seventh day. It
		is helpful to have a place to stay, wherever that may be.}
	{He returns after travelling a short distance. There is nothing to regret. There will be very good fortune.}
	{He returns resigned. There will be good fortune.}
	{He returns many times. This is dangerous. There is no mistake.}
	{He starts out in the midst of others, but returns alone.}
	{He returns with dignity. There is no need for regret.}
	{He has been given misleading information about his return. This is bad. There will be misfortune and
		mistakes. If he summons the army there will be a resounding defeat, which will also be disasterous
		for the country's ruler. It will not be possible to repel the invaders for another ten years.}
\or \interpret{WU WANG}{Not False}
	{Not false. Great success. It is good to remain firm. He will be distressed if he is wild instead of
		upright. It is useless for him to try to go in any direction.}
	{He is sincere. Wherever he goes he will meet good fortune.}
	{He does not plough the ground or clear the stable. But he still benefits from this wherever he goes.}
	{Even those who are truthful suffer misfortunes, like an ox which is tethered to a post and stolen by
		a passer-by. A local person suffers as a result of this misfortune.}
	{If he is able to stand firm there will be no mistakes.}
	{He is ill even though he is open and truthful. He does not use medicine. He will rejoice.}
	{He is not false yet if he tries to move he will be in difficulties. At this time it is wiser not to move.}
\or \interpret{TA CH'U}{Great Domesticating Powers}
	{Great powers of domestication. It is helpful to stand firm. He is lucky when he does not eat with his
		family. It is helpful to cross the great river.}
	{He is in a dangerous situation. It is wise to stop.}
	{The main strap underneath the carriage has been taken away.}
	{He urges his good horse to travel on. It will be helpful to realize the dangers and behave accordingly.
		He trains daily in chariot driving and self-defence. There will be advantages in whatever direction he
		takes.}
	{The wild young bull is restrained by a yoke. There will be good fortune.}
	{The tusks of the gelded boar. There will be good fortune.}
	{He is the one who controls the very substance of Heaven. There will be success.}
\or \interpret{I}{Taking Nourishment}
	{Taking nourishment. By behaving properly there will be good fortune. Look at what we are seeking to nourish.
		How people naturally seek nourishment.}
	{Let your sacred tortoise go and look at me with your mouth open. There will be misfortunes.}
	{He turns everything upside down for nourishment. It goes against the tradition of seeking
		nourishment above the ground. This will lead to evil.}
	{He rejects nourishment. This will be bad no matter how firm he is. He does nothing for ten
		years. This will not bring any disadvantages.}
	{It is good if he turns things upside down for nourishment. Like a tiger who glares down, he is ready
		to drive them out. This is no mistake.}
	{He is opposed to the correct traditional teachings. He will have good fortune if he lives properly.
		He should not be allowed to try to cross the great river.}
	{The source of nourishment. It is dangerous but there will be good fortune. It is helpful to
		cross the great river.}
\or \interpret{TA KUO}{Great Experience}
	{Great experience. The main support is weak. It is helpful to move in any direction. There will be success.}
	{He puts white woven mats on the ground. There will be no mistakes.}
	{The ageing willow sprouts new shoots. The old husband has married a young wife. There are no disadvantages.}
	{The main support is weak. This is bad.}
	{The main support has held. This is good fortune. Another support will only cause regret.}
	{The ageing willow flowers. The old wife marries a young husband. There wil be mistakes and no praise.}
	{He crosses the stream and the water rises above his head. This is bad, but there are no mistakes or blame.}
\or \interpret{K'AN}{The Watery Depths}
	{The watery depths are twofold. Have faith. He keeps his heart faithful and prospers. Whatever he does is
		worthwhile}
	{The waters are twofold. The man enters the watery cavern. This is evil.}
	{The watery depths are dangerous. His desires will not be fully satified.}
	{He is confronted by the watery depths whether he goes forward or retreats. When danger such as this is
		around he should stop and rest, otherwise he will fall into the watery depths. Do nothing.}
	{He only has a bottle of wine and a basket of rice and he uses ordinary earthenware pots. He has agreed
		to bring enlightenment. There is nothing wrong in this.}
	{The watery depths are not overflowing. Order is approaching. There will be no mistakes.}
	{He is bound with thick strong ropes and imprisoned behind walls of thorn bushes. For three yaers he cannot
		find a means of escape. This is evil.}
\or \interpret{LI}{To Shine Brightly, to Part}
	{To Part. It is useful to stand firm and behave well. This will bring success. Take care of the cows. There
		will be good fortune.}
	{He moves but in a very confused way. If this is done with good and proper intentions there will be no mistakes.}
	{The yellow parts. There will be good fortune.}
	{The setting sun shines as it goes down. The old sing and beat their pots, or complain about their lot.
		This is evil.}
	{It comes unexpectedly. It is like a fire which dies down and is discarded.}
	{He leaves in floods of tears, crying and bemoaning. There will be good fortune.}
	{The king uses him to attack. Skilfully he kills the leaders but spares the followers. This is not a mistake.}
\or \interpret{HSIEN}{All-Embracing}
	{All-embracing. It is helpful to stand firm and behave properly. Marriage to a young woman will be fortunate.}
	{He concentrates on wriggling his toes.}
	{He concentrates on the calves of his legs. This is evil. If he stays where he is there will be good fortune.}
	{He concentrates on flexing his thigh muscles. He follows others closely. Going forward in this way will
		bring unhappiness.}
	{By being firm and righteous there will be good fortune. This prevents regret. If he is restless, moving to
		and fro, only his friends will take him seriously.}
	{He concentrates on the back of his neck. There will be no regrets.}
	{He concentrates on his jaw and tongue.}
\or \interpret{HENG}{Constatnt}
	{Constant. There will be success. There is no mistake. It is helpful to stand firm and behave righteously.
		It is helpful to go forward in any direction.}
	{Too great an emphasis on constancy willl bring evil, even though he behaves firmly and righteously. Nothing
		at all can be gained here.}
	{All regret is removed.}
	{He does not always make use of virtue. Others think of this as a disgrace. Even if he behaves properly
		there will be mistakes.}
	{There is no game in the field.}
	{By righteous behaviour he increases his virtue. This is good for the wife but bad for the husband.}
	{He constantly excites himself. This is evil.}
\or \interpret{TUN}{To hide}
	{To hide. There will be success. There is a small advantage to be gained from being firm and behaving properly.}
	{It is dangerous because the tail is hidden. Do not try to move.}
	{He grasps it tightly, using an ox's yellow hide. He cannot be persuaded to let go.}
	{He hides away as thought he is bound to this place. He is ill and troubled. Keeping domestic servants or a
		concubine will bring good fortune.}
	{He hides away well. This is fortunate for a wise man. It is impossible for a lesser man.}
	{He hides himself away very carefully. Righteous behaviour will bring good fortune.}
	{He hides away honourably. This will bring many advantages.}
\or \interpret{TA CHUANG}{Great Strength}
	{Great strength. It is helpful to act properly and firmly.}
	{His strength lies in his toes. It is certainly a bad time to attack.}
	{We will be fortunate if he behaves properly.}
	{The small man uses all his strength, whereas the wise man does not do this. There is a grave danger even
		when he behaves properly. He is like a ram who butts a fence and only succeeds in trapping his horn.}
	{Righteous conduct will bring good fortune and remorse will disappear. He approaches a fence determined
			not to get trapped. His strength is like the strength of a heavy wagon's wheels.}
	{He changes and loses his ram-like nature. There is nothing to regret.}
	{He is like a ram butting a fence. He cannot retreat or advance. He cannot gain anything by carrying on like
		this. If he understands the dangers there will be good fortune.}
\or \interpret{CHIN}{To Advance}
	{To advance. The vigorous noble. He uses his charm like a display of fine horses. Three times within the
		space of one day he is received in audience.}
	{He seems to advance and retreat at the same time. If he behaves properly all will work out well. If
		confidence is expressed remain calm. There will be no mistakes.}
	{He appears to advance, but at the same time he is sad. If he behaves properly he will receive very good
		fortune from his heavenly mother.}
	{Everyone is in harmony. All regret has disappeared.}
	{He is like a hamster when he advances. He is in danger even if he behaves properly.}
	{All regrets have now gone. Do not be worried by what is gained or lost. To advance will bring good fortune
		and many advantages.}
	{He advances with his horns. He only uses them to punish the city. There will be good fortune. There is no
		mistake. Even if he behaves properly there will be problems.}
\or \interpret{MING I}{Brightness Dimmed.}
	{Brightness dimmed. It is wise to appreciate the dangers and behave properly.}
	{The brightness is dimmed with flying and he lowers his wings. The wise man goes away and does not eat
		for three days. Wherever he travels people speak of him unkindly.}
	{Brightness is dimmed. He is wounded in the left thigh. He uses a swift horse to save himself. There
		will be good fortune.}
	{While hunting in the south the brightness is dimmed. He captures the most important general. He should not
		be too hasty when he tries to restore order.}
	{He enters the left side of the belly. He reaches the heart of that which dims the brightness. He leaves
		through the courtyard gate.}
	{Wherever brightness is dimmed, be like Prince Chi. It is good to behave properly.}
	{No light, just obscurity. First of all he climbs to Heaven. Afterwards he returns to earth.}
\or \interpret{CHIA JEN}{The Family}
	{The family. It is good if a woman behaves properly.}
	{He has restrictions in his family. Regrets will disappear.}
	{She should not follow her own desires. She should be at the centre of the home preparing food. Her good
		conduct brings great fortune.}
	{Excessive severity in the family brings repentance and dangers. There is also good fortune. If the women and
		children are always chattering this will end in distress and humiliation.}
	{The wealthy family. There is extremely good fortune.}
	{The Emperer is very forgiving with his family. There is nothing to worry about. There will be good fortune.}
	{He is confident and has a regal presence. There will be good fortune in the end.}
\or \interpret{K'UEI}{Opposition}
	{Opposition. There will be good fortune in minor matters.}
	{There is no need for repentance. He has lost his horse but does not go in search of it because it will
		return of its own accord. If he meets evil men he should avoid making any mistakes.}
	{By chance he meets his lord in a narrow street. He does not make any mistakes.}
	{The wagon is seen as it is dragged backwards. His oxen are obstructed. The man's own hair and nose are
		cut off. This is not a good start, but the end will be better.}
	{Alone in opposition. He can be confident when he meets the top man. There is danger but there will be no
		mistakes.}
	{There is no need for repentance. His own relative bites through the covering. Where is the fault in
		carrying on like this?}
	{Alone in opposition. He can see the pig carrying all the filth on its back. There is a carriage full of
		ghosts. At first the bow is drawn, then it is unstrung. He is not a robber in the act of theft,
		but a bridegroom going to woo. If he carries on along this way he will encounter rain and there
		will be good fortune.}
\or \interpret{CHIEN}{Obstruction}
	{There are advadtages in the southwest. There are no advantages in the sortheast. It is helpful to see the
		great man. Behave properly and there will be good fortune.}
	{The way ahead is blocked. Returning will bring praise.}
	{The king's servant faces many obstructions and difficulties. He is not to be blamed.}
	{To go forward is to encounter difficulties. The reverse of this is to return.}
	{To go forward is to encounter obstacles. By returning he is united.}
	{There are very great obstacles. His friends come back to him.}
	{To go forward is to encounter obstacles. It is better if he turns back. There will be good fortune.
		It is helpful to see the great man.}
\or \interpret{HSIEH}{Let Loose}
	{Let loose. There are advantages in the southwest. If there is nothing else to be done there is good fortune in
		his returning. If he has to go anythere it will be more fortunate to travel early.}
	{There will be no mistakes.}
	{He catches three foxes in a field. He receives the yellow arrow. Proper behaviour brings good fortune.}
	{The man is carrying a burden on his back. He is also travelling in a carriage and this will lead to robbers
		attacking him. No matter how properly he behaves there will be sadness.}
	{Free yourself from your toe. Then your trusted friends will come to you.}
	{The wise man removes the carriage and has good fortune. He has the confidence of a lesser man.}
	{The lord fires an arrow at a falcon which is perched on top of a wall. He hits it. The act brings
		advantages to him.}
\or \interpret{SUN}{Injured}
	{Injured. Have confidence. Supreme good fortune and no regrets. Proper behaviour is possible. It is good to
		advance when fully prepared. Two baskets of offerings may be better than offerings which are
		expensive.}
	{He leaves his own work. He goes to help another person. There is no cause for regret if he deliberates
		on what he is doing.}
	{It is good to behave properly and to attack evil. It is not possible to help another without helping himself.}
	{Three man travel together, but one then disagrees with the others and he therefore goes. He then finds
		another person as a companion}
	{He tackles his weakness and is then able to help others. This is good. He has nothing to regret by doing this.}
	{He may receive advantages because friends give him ten tortoise shells. They will not accept them back.
		There will be very great fortune.}
	{He does not injure. He has the advantage and there is no cause for remorse. It is good to remain firm and
		true because good fortune will naturally arise. He has the advantage in every move. He draws to
		himself able ministers, not just relatives.}
\or \interpret{I}{Increase}
	{Increase. It is helpful to go forward with plans in mind. It is good to cross the great river.}
	{It is good to undertake impressive actions. There will be great fortune and nothing to regret.}
	{Someone gives him the advantage by giving him ten pairs of tortoise shells. These cannot be resisted.
		He will always be fortunate as long as he behaves properly. The king who offers what he receives
		to the supreme ruler will be very fortunate.}
	{He gains advantage by using evil. There is nothing wrong in this. He has confidence. He walks the
		straight path and, using the seal, announces himself to the prince.}
	{He walks the true path. He announces himself to the prince who will follow. It is good to be relied upon,
		even to the extent of helping to remove the capital.}
	{He is confident in his generosity and kindness. He can be confident because he will recognise his virtue
		and kindness}
	{He is of no benefit to anyone. Someone may even hit him. He sets no standard for his attitudes and this
		is evil.}
\or \interpret{KUAI}{New Outcome}
	{New outcome. He must display it in the king's court. He must speak confidently and with strict regard
		to honesty. There is danger. It is not a good idea to tell the city, nor is it useful to rush to
		arms. It is helpful to carry on with plans already in action.}
	{He steps out boldly. He sets forth, but not to victory. He will regret this.}
	{He stands almost ready whilst others arm in the night. There is nothing to fear.}
	{He puts his best face forward and this brings misfortune. The wise man is securely positioned. He walks
		above and encounters rain, and is therefore soaked and annoyed. There is no cause for regret.}
	{The skin on his buttocks has been stripped away and he walks with difficulty. If he were led like a sheep
		there would be no worries. However, if he were to hear what is said he would not believe it.}
	{When he sorts out weeds he is secure in his determined attitude. There will be no mistake in following
		the correct path.}
	{He does not stand prepared. This will eventually bring evil.}
\or \interpret{KOU}{To Meet}
	{To meet. The woman is strong and tough. Do not get married to a woman like this.}
	{He should be restrained with an iron brake. Proper behaviour will bring good fortune. If he goes forward
		in any way he will encounter misfortune. He is like a thin pig who rushes around.}
	{There is nothing wrong in the fact that he has a packet of fish. Do not give this to guests.}
	{The skin of his buttocks has been stripped away and he has difficulty walking. There is danger but he
		does not make any major mistakes.}
	{He has a packet but no fish. There will be misfortune.}
	{The melon is covered by the tree's leaves. Part of it is hidden. It comes as though from Heaven.}
	{He meets others with his horns. This will bring regret. There will be no mistakes.}
\or \interpret{TS'UI}{To Collect}
	{To collect. Success. The king approaches the temple. It is good to see the great man. There will be
		success. The use of large offerings brings good fortune. To move forward in any direction
		will also bring good fortune.}
	{He is sincere in his wish, but if he does not carry it out there may sometimes be confusion, sometimes
		a coming together. He calls out. Then his tears give way to smiles. He does not mind. He goes
		forward and there will be no mistakes.}
	{He is led forward and there is great fortune. There are no mistakes. As long as he is sincere even the
		smallest offering will bring advantages.}
	{He tries to bring unity, but then he sighs. He can find no advantages anywhere. There will be no
		mistakes if he goes forward, but there may be slight cause for regret.}
	{Great fortune. There will be no mistakes.}
	{He uses his dignified position to unite them together. This is no mistake. If someone is insincere,
		then admirable, persistent and good behaviour will ensure that distress disappears.}
	{He sighs and weeps floods of tears. There will be no mistakes.}
\or \interpret{SHENG}{Rising Up}
	{Rising up. There will be great success. He tries to see the great man. He does not worry. Moving to
		the south will bring good fortune.}
	{He rises up and is welcomed. There will be extremely good fortune.}
	{If he is sincere then even the smallest offerings will be advantageous. There will be no mistakes.}
	{He goes up into an empty town.}
	{The king uses him to make offerings for success on Mount Ch'i. There will be good fortune. There
		will be no mistakes.}
	{He behaves properly and is fortunate. He mounts the steps.}
	{He rises up from the dark. It is helpful to be unceasingly persistent.}
\or \interpret{K'UN}{To Surround and Wear Out}
	{To surround and wear out. There will be success. Keep going. The great man has good fortune. There
		will be no mistakes. He has a speech to make but he is not believed.}
	{He is sitting exhausted under the barren trees. He wanders into a dreary valley. Nothing happens
		for three years.}
	{He is sitting exhausted amid his wine and meat. The person with the red knee bands is coming. It is
		helpful to offer sacrifices. Moving forwards brings bad fortune but there will be no mistakes.}
	{He is sitting exhausted among the stones. He grasps at thorns. He goes into his palace and does not
		see his wife. There will be bad fortune.}
	{He proceeds very slowly. He is sitting exhausted by his golden carriage. There is good reason for his
		sadness but all will be well in the end.}
	{His nose and feet are cut off. He is sitting, exhausted by the person with the scarlet knee bands.
		Everything is slow but well. It is helpful to make sacrifices.}
	{He is sitting exhausted among the creeping plants. He moves in a certain way and says ``If I move I
		shall be sorry.'' If he is able to be sorry and then moves forward there will be good fortune.}
\or \interpret{CHING}{The Well}
	{The well. The town may be moved, but the well may not. It does not shrink or grow. People come and go
		to draw water from the well. If they cannot quite reach the water with a rope or if the
		bucket breaks, there will be bad fortune.}
	{He does not drink from the muddy well. No creatures come to an old well.}
	{The well has sprung a leak and the fish and water spurt forth. The broken jug leaks.}
	{Although the well has been cleaned no one uses it. This is a weight upon the heart because it can
		be used. If the king were clever he and others could benefit from this.}
	{The well is well lined. There will be no mistakes.}
	{The well has clear, cold water which can be drunk.}
	{He draws from the well without difficulty because it is open. He is sincere. This will bring great good
		fortune.}
\or \interpret{KO}{Change}
	{Change. He is believed in his own right time. There will be great success. It is helpful to behave
		properly. Any cause for regret disappears.}
	{He is bound in the skin of a yellow cow.}
	{When his time comes he is able to make changes. Initiating action will bring extremely good fortune.
		There will be no mistakes.}
	{Initiating action will bring bad fortune. Even if he behaves properly there is danger. If he wishes to
		change things he will need to discuss them three times, and then people will have confidence
		in him.}
	{The need to repent disappears. He is believed. He alters the way things are and is fortunate.}
	{The great man changes into a tiger. Even though he has not yet consulted the oracle people have
		confidence in him.}
	{The wise man changes like a leopard, whereas the lesser man simple recomposes his face. To advance
		will bring bad fortune. To continue behaving correctly will bring good fortune.}
\or \interpret{TING}{The Cooking Pot}
	{Ting, the tree-legged cooking pot. Great fortune and sucess.}
	{The cooking pot is upturned and its feet are in the air. This is helpful for getting rid of useless
		stuff. A concubine is useful because of the son she has. There will be no mistakes.}
	{The cooking pot has food inside it. My colleagues are jealous but they cannot harm me. There will be
		good fortune.}
	{The handles on the cooking pot have been changed. He is halted in his progress. The pheasant's juicy
		meat is not eaten. The rain comes and sadness ends. In the end there will be good fortune.}
	{The legs of the cooking pot are broken. The noble's food is split and his clothes are soiled.
		There will be bad fortune.}
	{The cooking pot has yellow ears and gold handles. It is helpful to continue behaving properly.}
	{The cooking pot has rings of jade. There will be very good fortune. Everything here works to his advantage.}
\or \interpret{CHEN}{Shock}
	{Shock. Success. The shock comes but he looks out and is seen with cheerful words on his lips. Although
		the shock terrifies everyone for an hundred miles around, he does not drop the sacrificial
		spoon and cup.}
	{The shock comes and he looks around carefully, but then cheerful words come to his lips. There will be
		good fortune.}
	{He is in danger when the shock comes. He loses his treasure. He climbs nine hills. He does not go in search
	of his treasures but finds them again after seven days.}
	{The shock makes him very upset and disturbed. The shock makes him act but there is nothing wrong in this.}
	{The shock makes him as sluggish as mud.}
	{He comes and goes and the shock threatens danger. However, there is no problem because he has things to do.}
	{The shock causes him to be dismayed and he looks around anxiously. Initiating action will bring good
	fortune. He will have acted rightly if he moves before the shock reaches him although is has reached his
	neigbourhood. His family, however, may grumble about him.}
\or \interpret{KEN}{Resting}
	{Resting. Resting his back, he is not able to feel his body. He goes into the courtyard. He does not
	see his own people. There will be no mistakes.}
	{There is nothing wrong with resting his toes. It is helpful for him to behave in a true manner.}
	{He rests his legs. He is unable to save the person whom he follows and is uneasy in his heart.}
	{He rests his loins. He puts his ribs aside from him. There is danger and his heart bursts with
		anxious excitement.}
	{He rests his body. There will be no mistakes.}
	{He rests his jaws. His speech is composed. Any cause for regret disappears.}
	{He remains resting with devotion. There will be good fortune.}
\or \interpret{CHIEN}{Gradual Development}
	{Gradual development. The woman is married. There will be good fortune. It is helpful to behave properly.}
	{The geese gradually draw near to the shore. The younger son is in danger. Although he is talked about
		there will be no mistakes.}
	{The geese gradually draw near to the cliff, and here they eat and drink happily and contentedly. There
		will be good fortune.}
	{The geese gradually draw near to the plains. This is like a husband who goes out but does not return,
		or like a wife who is pregnant but does not have the baby. There will be bad fortune. It is good
		to fight off thieves.}
	{The geese gradually draw near to the tree and land on its flat branches. There will be no mistakes.}
	{The geese gradually draw near the mountain. This is like a wife who does not become pregnant for three
		years. In the end she conceives as nothing can prevent this. There will be good fortune.}
	{The geese gradually draw near to the highest heights. Their feathers can be used for the ritual dance.
		There will be good fortune.}
\or \interpret{KUEI MEI}{Marrying the Younger Sister}
	{Marrying the younger sister. Initiating actions will bring bad fortune. There are no advantages here.}
	{The younger sister is married as a subordinate to the first wife. This is like a lame man who manages
		to get along. To advance will bring good fortune.}
	{The one-eyed person is able to see. It is helpful for a man to behave in the same way as a woman does.}
	{The younger sister is ready to be married but she is in a lowly position. When she returns she is
		married off as a concibine.}
	{The younger sister spins out the time to her marraige, but she will eventually be married.}
	{Emperor I gave his younger sister in marriage but her embroidered clothes were not as gorgeous as
		those of her attendants. The moon is almost full. There will be good fortune.}
	{The woman carries a basket but it does not contain any food. The gentleman stabs the sheep but no
		blood comes out. There is nothing to be gained here.}
\or \interpret{FENG}{Prosperity}
	{Prosperity. Success. There is nothing to worry about because the king achieves everything. He should
		be like the sun at midday.}
	{He meets his supreme ruler. There is nothing wrong in the fact that they are alike. Going forward
		will result in praise.}
	{His properity is like a curtain. At midday he can see the stars in the Dipper constellation. When he goes
		out he is treated with suspicion and hatred. There will be fortune if he is encouraged to
		follow confidently}
	{His prosperity is like a banner through which he can see the smallest stars at midday. He breaks his
		right arm but there will be no mistakes.}
	{His properity is like a curtain, through which he can see the stars in the Dipper constellation at
		midday. He meets his king and they are alike. There will be good fortune.}
	{The various parts come together. He is praised and congratulated. There will be good fortune.}
	{His own household is prosperous. He encloses his family inside and peers out through the gate, but does
		not see anyone there. He does not see anything for three years. There will be misfortune.}
\or \interpret{LU}{The Traveller}
	{The traveller. Minor success. If the traveller behaves properly there will be good fortune.}
	{Because the traveller is engaged in worthless activities he brings disaster upon himself.}
	{The traveller arrives at an inn, carrying his goods. He obtains the services of a dependable and
		upright servant.}
	{The traveller burns down the inn. He loses his faithful servant. Even if he behaves properly there
		in danger.}
	{The traveller is in a shelter and has his goods and an ox with him. ``My heart is not at ease.''}
	{He shoots at a pheasant but loses his arrow. Eventually he is praised and given an official post.}
	{He is like a bird which burns its own nest. Initially the traveller laughs, but then he weeps. Because
		of the changes he loses his ox. There will be bad fortune.}
\or \interpret{SUN}{Gentle and Yielding}
	{Gentle and yielding. Minor success. It is helpful to advance. It is also helpful to see the great man.}
	{He advances and retreats. It is helpful to be well prepared and to behave like a soldier.}
	{He sits quietly under the bed. He uses diviners and magicians, causing great confusion. There
		will be good fortune and no mistakes.}
	{He is persistent but gentle. This is regrettable.}
	{There are no regrets. He catches three kinds of creatures in the field.}
	{Proper behaviour brings good fortune. There is nothing to regret. Everything he does is advantageous.
		The start may not be good but everything will work out well in the end. There will be three days
		before the changes occur. Allow three days after the changes. There will be good fortune.}
	{He sits quietly under the bed. He loses his goods and his ox. However well he behaves there will
		be misfortune.}
\or \interpret{TUI}{Happiness}
	{Happiness. Success. It is advantageous to behave properly and firmly.}
	{Happiness and harmony. There will be good fortune.}
	{Confidence and happiness. There will be good fortune. Every sadness disappears.}
	{Happiness is coming. There will be misfortune.}
	{He deliberates about what will make him happy but this does not satisfy him. The things he is involved
		with are diseased but he will be happy.}
	{He is confident despite the disasters around him. This is dangerous.}
	{Seductive happiness.}
\or \interpret{HUAN}{Scattered}
	{Scattered. Success. The king travels to his temple. It is helpful to cross the great river. It is also
		good to behave properly.}
	{He rescues with the strength of a horse. There will be good fortune.}
	{When everything is scattered abroad he hurries to a place of safety. There is nothing to be
		anxious about.}
	{He disregards his own person. This is no cause for regret.}
	{He breaks up his group. There will be great fortune. He scatters and then gathers in a mound;
		ordinary people would never have considered doing this.}
	{He is dripping with sweat when he loudly cries out. He has done nothing wrong. He scatters the king's
		officers. There is nothing wrong in this.}
	{He scatters his blood. He departs. He leaves and then keeps his distance. There will be no mistakes.}
\or \interpret{CHIEH}{Limitations}
	{Limitations. Success. If the limitations are too strict they should not be allowed to go on for too long.}
	{He does not go out of the courtyard outside his door. There will be no mistakes.}
	{He does not go out of the courtyard inside his gate. There will be bad fortune.}
	{Tears will fall because he does not observe the restrictions. There is no fault in this.}
	{He accepts the limitations peacefully. There will be success.}
	{He willingly accepts the limitations. There will be good fortune. His advance is praised.}
	{He is distressed by the limitations. Even when he acts properly there will be misfortune. Although
		there are regrets, they will disappear.}
\or \interpret{CHUNG FU}{Inner Confidence}
	{Inner confidence. Pigs and fish mean good fortune. It is helpful to cross the great river. It is
		helpful to behave properly.}
	{He is ready. There will be good fortune. Otherwise he would be uneasy.}
	{The crane calls out from her cover. Her chicks answer back. ``I have a good cup full which I will
		share with you.''}
	{He meets his companions. He beats the drum amd then stops. He weeps and then sings.}
	{The moon is almost full. The horse has lost his mate. There is nothing wrong in this.}
	{He is united with others when he is confident. There will be no mistakes.}
	{The voice of the cockerel rises up to Heaven. Even if be behaves properly there is bad fortune.}
\or \interpret{HSAIO KUO}{Minor Problems}
	{Minor problems. Success. It is useful to behave properly. This is appropriate for minor things but not
		for more important things. The bird on the wing brings the message. It is good to come down but
		not to go up. There will be extremely good fortune.}
	{The bird is on the wing. There will be misfortune.}
	{He passes by his ancestor. He meets his grandmother. He does not do anything against his ruler.
		He meets his official. There will be no mistakes.}
	{Someone comes up and attacks him because he is not sufficiently careful. There will be misfortune.}
	{There are no mistakes here. Instead of passing by he meets him. Carrying on brings danger and care
		should be taken. Do noy always try to use proper behaviour.}
	{There is a heavy cloud which is coming from our lands in the west, but no rain. The prince fires and
		hits the bird in the cave.}
	{He does not meet him but passes by. The bird on the wing has flown far away from him. There will be
		bad fortune. This is a sign of pain and misfortune.}
\or \interpret{CHI CHI}{Already Done}
	{Already done. Minor success. It is helpful to behave properly. There will be good fortune to begin with,
		but chaos at the end.}
	{He puts the brakes on his wheels. He is like a creature who has wet his tail. There will be no mistakes.}
	{He is like a woman who has lost her screen. She does not go after it. She will get it back in seven days.}
	{The noble ancestor destroyed the devil's territory but took three years to subdue it. Lesser people
		should not be used.}
	{The best clothes are now in rags. In the end he will be careful all day.}
	{His neighbour is the east kills an ox, but he does not have as much as his neighbour in the west who
		sacrifices. His sacrifice brings great blessings.}
	{His head is in the water. There is danger.}
\or \interpret{WEI CHI}{Not Yet Done}
	{Not yet done. Success. A young fox crossing the stream gets his tail wet. There is no advantage in this.}
	{He gets his tail wet. This will cause regret.}
	{He puts the brakes on his wheels. Proper behaviour will bring good fortune.}
	{He advances and meets misfortunes before things are properly finished. It is helpful to cross
		the great stream.}
	{Proper conduct will bring good fortune. All regret passes away. He is shaken and invades the devil's
		territory. The great country rewards him for three years.}
	{Proper behaviour will bring good fortune. There is no cause for regret. He has the glory of the
		wise man and he is sincere. There will be good fortune.}
	{He is confident and celebrates with drink. There is no wrong in this. If he wets his head he will
		lose his confidence.}
\fi
\end{document}
